---
title: "chapter 3"
author: "Nick Marks"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
devtools::install_github("rmcelreath/rethinking")
library(rethinking)
library(tidyverse)
options(scipen = 9)
# data(homeworkch2)
```



```{r pg40}
#Define grid
p_grid <- seq(from=0,to=1, length.out = 1000)

#define prior
prior <- rep(1,1000)

#compute likelihood at each value in grid
likelihood <- dbinom(6, size=9, prob=p_grid)

#compute product of likelihood and prior
unstd.posterior <- likelihood * prior


#standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

#sample posterior
sample <- sample(p_grid, prob = posterior, size = 10000, replace=TRUE)
dens(sample)
plot(sample)

#probabilty world water proportion is less than 50%
sum(posterior[p_grid < .5])
sum(sample < .5) / 10000
```


# Easy practice
```{r chapter 3 practice}
p_grid <- seq(from=0,to=1,length.out = 1000)
prior <- rep(1,1000)
likelihood <- dbinom(6, size=9, prob=p_grid)
posterior <- likelihood*prior
posterior <- posterior/sum(posterior)
set.seed(100)
samples <- sample(p_grid, prob=posterior, size=10000, replace=TRUE)
```

How much posterior proability lies below p = 0.2? p = 0.8? p between .2 and .8?
```{r 3E1}
sum(posterior[p_grid < .2])
sum(samples < 0.2) / 1e4
sum(posterior[p_grid < .8])
sum(samples < 0.8) / 1e4
sum(posterior[p_grid < .8 & p_grid > .2])
sum(samples > 0.2 & samples < 0.8) / 1e4
```

20% of the post prob lies below which value of p? above which value of p? min 66% interval? central 66% interval?
```{r}
quantile(samples, .2)
quantile(samples, .8) - quantile(samples, .2)
HPDI(samples, prob = .66)
PI(samples, prob = .66)
```


#Medium practice
8 water on 15 toses scenario
```{r}
p_grid <- seq(from=0,to=1,length.out = 1000)
prior <- rep(1,1000)
likelihood <- dbinom(8, size=15, prob=p_grid)
posterior <- prior*likelihood
posterior <- posterior / sum(posterior)
samples <- sample(p_grid, prob=posterior, size = 1e4, replace=TRUE)
HPDI(samples, p=.9)
```

Simulate parameter uncertainty for each value of samples
```{r}
w <- rbinom(1e4, size=15, prob=samples)
simplehist(w)
```

#Hard
```{r load data}
data(homeworkch3)
boy_births <- sum(birth1) + sum(birth2)
births <- data_frame(boy = c(birth1, birth2)
                     , birth_order = c(rep(1,length(birth1))
                                       , rep(2,length(birth2))))
```

#### 1a. Posterior distribution of birth = boy (1)
```{r}
sample_size <- 10000
model <- data_frame(p_grid = seq(0,1,length.out=sample_size)
                    , prior = rep(1, sample_size))

model <- model %>% 
  mutate(likelihood = dbinom(sum(births$boy), size = nrow(births), prob = p_grid)
         , posterior = prior*likelihood / sum(prior*likelihood))

ggplot(model, aes(p_grid, posterior)) +
  geom_line()

p_grid[which.max(posterior)]
```


#### 1b. 50%, 89%, and 97% posterior density intervals
```{r}
model <- model %>% 
  mutate(sample_grid = sample(p_grid, size = sample_size, replace = TRUE, prob = posterior))

samples_hpdi <- HPDI(model$sample_grid, prob = c(.5,.90,.95))
samples_hpdi

model <- model %>% 
  mutate(sample_cat = case_when(.$sample_grid >= samples_hpdi[3] & .$sample_grid <= samples_hpdi[4] ~ '50%'
                                , .$sample_grid >= samples_hpdi[2] & .$sample_grid <= samples_hpdi[5] ~ '90%'
                                , .$sample_grid >= samples_hpdi[1] & .$sample_grid <= samples_hpdi[6] ~ '95%'
                                , TRUE ~ 'Outside'))

ggplot(model, aes(sample_grid, fill = sample_cat)) +
  geom_histogram(bins = 1000) +
  scale_x_continuous(limits = c(0,1))
```

#### 1c.
```{r}
set.seed(20170428)
boys_qa <- rethinking::map(
    alist(
        boys ~ dbinom(size=nrow(births),prob = pr.boy) ,  # binomial likelihood
        pr.boy ~ dunif(0,1)    # uniform prior
), data=list(boys = sum(births$boy)) )

boys_mean <- precis(boys_qa)@output$Mean
boys_sd <- precis(boys_qa)@output$StdDev

set.seed(20170428)
model <- model %>% 
  mutate(sample_qa = rnorm(n=sample_size, mean=boys_mean, sd=boys_sd))

ggplot(model,(aes(x=samples_qa))) +
  geom_histogram(bins=1000) +
  scale_x_continuous(limits=c(0,1))

ggplot(model) +
  geom_density(aes(x=sample_grid, color = "blue")) +
  geom_density(aes(x=sample_qa, color = "red")) +
  scale_color_manual(values = c("red","blue")
                     , labels = c("Grid","Quadratic"))


```

#### 2a. Predict all births
```{r}
set.seed(20170428)
model <- model %>% 
  mutate(bsim = rbinom(sample_size, nrow(births), prob = sample_qa))

ggplot(model, aes(x=bsim)) +
  geom_density() +
  geom_vline(xintercept = sum(births$boy), color = "red")

```


#### 2b. Predict first born boys
```{r}
set.seed(20170428)
model <- model %>% 
  mutate(b1sim = rbinom(sample_size, nrow(births %>% filter(birth_order == 1)), prob = sample_qa))

ggplot(model, aes(x=b1sim)) +
  geom_density() +
  geom_vline(xintercept = sum(births$boy[births$birth_order==1]), color = "red")

```

#### 2b. Predict second borns with sisters
```{r}
b01 <- birth2[birth1==0]
set.seed(20170428)  
model <- model %>% 
  mutate(b01sim = rbinom(sample_size, length(b01), prob = sample_qa))

ggplot(model, aes(x=b01sim)) +
  geom_density() +
  geom_vline(xintercept = sum(b01), color = "red")

```
