---
# title: Template Report
page_header: true
header_title: Week 2 Homework
legal: "&copy; 2017 All rights reserved. Proprietary and confidential, Elicit LLC."
output:
  bookdown::tufte_html2:
    number_sections: false
    toc: false
    template: elicit_template.html
css: elicit_report_template.css
---

```{r setup, include=FALSE}
#NOTE: bookdown and tufte packages must be installed to knit this template
# install.packages('bookdown')
# install.packages('tufte')

pacman::p_load(tufte, knitr, tidyverse, lubridate, plotly, scales, 
               xtable, magrittr, rstudioapi, stringr, Hmisc, eboxr) 
options(scipen = 9)

# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'), 
                      echo=FALSE, cache = TRUE, comment = NULL,
                      dev = "cairo_pdf")
options(htmltools.dir.version = FALSE)
old <- theme_get()
the_new <- old %+% theme(text = element_text(family = "HelveticaNeue-Light",
                                             size = 10),
    panel.background = element_rect(fill = "transparent",colour = NA),
  panel.grid.major = element_line(colour = "grey"))
theme_set(the_new)
options(xtable.comment = FALSE)
options(xtable.booktabs = TRUE)
lpercent <- function(x) Hmisc::latexTranslate(percent(x))
ldollar <- function(x) Hmisc::latexTranslate(dollar(x))
discount_rate <- (1.12)^(3/2)-1
xtable <- function(x, ...) {
  for (i in which(sapply(x, function(y) !all(is.na(match(c("POSIXct","Date"),class(y))))))) x[[i]] <- as.character(x[[i]])
  xtable::xtable(x, ...)
}
cap_all_words <- function(name) {sapply(name, function(x) {
  paste(sapply(strsplit(x, ' '), capitalize), collapse=' ')
})}
rprint <- function(tbldf,dollars=c(),xtable_args = list(), NA.string="NA", ...) {
  if(rstudioapi::isAvailable()) {return(print(tbldf))}
  xt <- do.call(xtable, c(list(x=tbldf), xtable_args))
  if(length(dollars)>0) colnames(xt)[dollars] <- paste(colnames(xt)[dollars],"($)")
  colnames(xt) <- cap_all_words(str_replace_all(colnames(xt),'_',' '))
  print(xt,include.rownames=FALSE,NA.string=NA.string,
      type='latex',
      format.args=list(big.mark=","), ...)
  invisible()
}
```

```{r load data, include = FALSE}
library(rethinking)
data(homeworkch3)
sample_size = 10000
```

# 1. Back to births
## (a) Compute p-value for null hypothese that the probability of a boy birth is 0.5

There were 111 male births out of 200 observations. The null hypothesis of p = 0.5 would correspond to 100 boy birhts. Assuming this is true, we  want to calculate the area in tail of the null distribution that's beyond this point estimate (i.e. >111)
```{r 1a}
p <- dbinom(x=111:200, size = 200, prob=.5)
ggplot(NULL, aes(111:200, y=p)) + 
  geom_point()
```
We find that almost 7% of the area under the curve remains in the tail, so using the standard 5% p-value cutoff we would not reject the hypothesis that p = 0.5 
```{r}
sum(p)
```

## (b) how does p-value compare to bayesian analysis from last week?
```{r}
set.seed(20170428)
boys_qa <- rethinking::map(
    alist(
        boys ~ dbinom(size=nrow(births),prob = pr.boy) ,  # binomial likelihood
        pr.boy ~ dunif(0,1)    # uniform prior
), data=list(boys = sum(births$boy)) )

boys_mean <- precis(boys_qa)@output$Mean
boys_sd <- precis(boys_qa)@output$StdDev

precis(boys_qa)@output

d <- rnorm(sample_size, mean = boys_mean, sd = boys_sd)
hist(d)

length(d[d >= 111/200]) / length(d)
```

# 2. Casino scenario: when Pr(Ho|D) != P
## (a) p-value of null [a fair die]
```{r}
p <- dbinom(x=18:65,size=65,p=1/6)
plot(p)
sum(p)

```
Based on the p-vale of 0.0175 it's highly unlikely the die is fair. But....

##(b) bayesian calculation of fairness (grid approach)
```{r}
model <- seq(0,1,length.out = 100)
prior <- rep(.5,100)
likelihood <- dbinom(x=18, size=65, prob=model)
posterior <- prior*likelihood / sum(prior*likelihood)
ggplot(NULL, aes(model, posterior)) + geom_point()
```
Here we are given additional information that a loaded die will come up as 1 50% of the time. Given this information, it's far more likely the die is NOT loaded than it is, as the MLE is around 25%, which is much closer to 1/6 than 1/2



<!-- Bottom gutter to keep content above footer  --> 
<div id="bottom-gutter" style="height: 60px;"></div>
